### 2020-03-30 05-translation.tex
fino a sect4 e flat
update fino a sezione 4 inclusa e def di flat

\paragraph{Processes for entities.}
Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\hat{s}_i$, $\hat{s}_o$ are used to test the provisioning of $s$ from the context;
names $\tilde{s}_i$, $\tilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ from the context; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ in the system.
We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ be the one for its absence.
They can be seen as instances of the same template, which is given below.

\[
\begin{array}{lcl@{\hskip 1cm}lcl}
 P_s &\defeq & P(s,\tilde{s},\hat{s},\underline{s}) 
 &
 \overline{P_s}& \defeq & P(\overline{s},\tilde{s},\hat{s},\underline{s})
\end{array}
\]
%{\color{red}
%\[
%\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h_j,k_j\in\{0,1\}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j}\ \link{\hat{s}_i}{\hat{s}_o}.P_s\\
%&&\ + \\
%&&
%\sum_{\scriptsize \begin{array}{l}h_j,k_j, \\ h'_j\in\{0,1\}\end{array}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j} \  \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} \    (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h'_j} \ \link{\underline{s}_i}{\underline{s}_o}.P_s \\
%
%%\sum_{R_1+R_2\geq 0}(\sum_{h,k \in[0,1]} 
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_1} \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} (\sum_{h ,k\in[0,1]}
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_2} \link{\underline{s}_i}{\underline{s}_o}.P_s\\
%&&\ +\\
%&&\sum_{j\geq 0}(
%  \startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^j\ \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
%\end{array}
%\]
%}
\[
\begin{array}{rcl}
 P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\end{array}
\]
The first line of $ P(s,\tilde{s},\hat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\hat{s}_i}{\hat{s}_o}$).
Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
values of $h$ and $k$.
The second line accounts for the analogous case where $s$ is not provided by the context ($\link{\underline{s}_i}{\underline{s}_o}$). 
The condition $k\geq 1$ guarantees that $s$ will remain present (the continuation is $P_s$).
%Here taking
%$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
The third line accounts for the case where $s$ is tested for presence, but it is neither produced nor provided by  the context. Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence.
%Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
% permutations of the solid links; nevertheless if the reactions, and their enumeration, is known
% the code of $P_s$ can be drastically simplified.

\paragraph{Processes for reactions.}
We assume that each reaction $a$ is assigned a progressive number $j$.
The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\tilde{c}_i}{\tilde{c}_o}$ is requested for any $c\in P_j$).
The next case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
The process is recursive so that reactions can be applied at any step.
\[
\begin{array}{lcl@{\quad}l}
P_{aj}& \defeq & 
 ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
&& + \ \\
&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
&&+\\
&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
\end{array}
\]

\noindent
We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $aj$.
Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using $cxt_j$ channels, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels). 
In the following there is an example explaining this schema.
%\[
%\begin{array}{lcl@{\qquad}l}
%P_{aj}& \defeq & 
% ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \link{}{}
% \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{r_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
%&& + \ \\
%&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainend{r_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
%&&+\\
%&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainend{r_{i+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
%\end{array}
%\]

\paragraph{Processes for contexts.}
For each entity $s \in S$, we introduce another process $Cxt_s$,  participating in each transition and saying if, the entity $s$ is provided by the context or not.
As done for the reactions, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
the chains formed by the application of all the contexts. 
For each entity $s$ with number $j$, at step $n >0$  there are two possible behaviours: 
\[
\begin{array}{lcl}
Cxt^n &\defeq &
%\left \{
\begin{array}{ll}
\sum _{j}^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in Ij_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in Oj_n} \,  \backslash_{p_1}. Cxt^{n+1}\\
%+ \\
% ^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in I2_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in Oj_n} \, \backslash_{p_1}
% \chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{p_1}. Cxt_{s}^{n+1} & \ \mbox{if the context provides $s$  at the $n$-th step}\\[5pt] 
% \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_s^{n+1} & \ \mbox{otherwise}
%\end{array}\right.\\
%Cxt_s & \defeq & Cxt_s^1\\
\end{array}
%\right.\\
\end{array}
\]

@@

We only consider $Cxt^n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (process $P_s$ instead of $\overline{P_s}$).

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
\begin{center}$\llbracket{\cal A},\gamma \rrbracket = \restrict{reacts,ctxs,ents,prods}(\Pi_{s\in C_0}P_{s} | \Pi_{s \notin C_0}\overline{P_s} | \Pi_{a\in A} P_a| \Pi_{s\in S} Cxt_s)$,
\end{center} 
with  $reacts$ be the set of reaction names $r_j $, $cxts$ the set of context names $cxt_j$,  $ents$ the set of decorated entity names  $\{s_i,s_o,\hat{s}_i,\hat{s}_o,\tilde{s}_i,\tilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $prods$ be the set of names $p_j$ associated to each reaction. 
In the following, we set $names = reacts\, \cup\, ctxs\, \cup\, ents\, \cup \, prods$.
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = p_1$ $p_{u+1}=\silent$ for $w$ the number of entities.
\end{definition}

@@

It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes $P_a$, with $a \in A$, and  $Cxt_s$ and  $P_s$, with $s \in S$,  be interacting in that transition.
This is due to the fact that  all the channels $r_j$, $p_j$, $cxt_h$, and $s_{hi}$, and $s_{ho}$ are  restricted.
Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation. If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{names}\upsilon} P')$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s_h \in S$ (where $h$ is the identifying number of $s$), the corresponding channel $s_h$ (suitably decorated), and the corresponding channel $cxt_h$ appear in $\upsilon$; 
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $Cxt_s$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.

@@

With the next proposition, we analyse the structure of a c\CNA \ 
process encoding of  a reactive process after one step transition.
In the following four statements, for brevity, we let ${\cal A} = (S, A)$ be a RS, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with $\gamma=\{C_i\}_{i\in\mathbb{N}}$ and $\delta=\{D_i\}_{i\in\mathbb{N}}$. Moreover, we denote by $\pi^j$ the shift of $\pi$ starting at the $j$-th state sequence; formally we let $\pi^j=(\gamma^j,\delta^j)$ with
$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ with 
  $C_0'=C_j \cup D_j$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.

 \begin{restatable}[Correctness 1]{proposition}{propcorruno}
% \begin{proposition}(correctness 1)\\
 \label{prop:corr1}
% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
 Let 
 $P=\llbracket{\cal A},\gamma\rrbracket$ with\\
$P  = \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$.\\
 If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:\\
  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$ and \\
 $P'= \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.\\
Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
 %\end{proposition}
\end{restatable}

@@

\noindent
With the following propositions, we prove that, given a RS ${\cal A} = (S, A)$ and an extended  interactive process $\pi=(\gamma,\delta)$, then the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of $\pi$.

>>>>

\paragraph{Processes for entities.}
Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\widehat{s}_i$, $\widehat{s}_o$ are used to test the provisioning of $s$ from the context;
names $\widetilde{s}_i$, $\widetilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ in the system; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ from the context.
We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ be the one for its absence.
They can be seen as instances of the same template, which is given below.

\[
\begin{array}{lcl@{\hskip 1cm}lcl}
 P_s &\defeq & P(s,\widetilde{s},\widehat{s},\underline{s}) 
 &
 \overline{P_s}& \defeq & P(\overline{s},\widetilde{s},\widehat{s},\underline{s})
\end{array}
\]
%{\color{red}
%\[
%\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h_j,k_j\in\{0,1\}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j}\ \link{\hat{s}_i}{\hat{s}_o}.P_s\\
%&&\ + \\
%&&
%\sum_{\scriptsize \begin{array}{l}h_j,k_j, \\ h'_j\in\{0,1\}\end{array}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j} \  \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} \    (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h'_j} \ \link{\underline{s}_i}{\underline{s}_o}.P_s \\
%
%%\sum_{R_1+R_2\geq 0}(\sum_{h,k \in[0,1]} 
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_1} \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} (\sum_{h ,k\in[0,1]}
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_2} \link{\underline{s}_i}{\underline{s}_o}.P_s\\
%&&\ +\\
%&&\sum_{j\geq 0}(
%  \startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^j\ \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
%\end{array}
%\]
%}
\[
\begin{array}{rcl}
 P(s,\widetilde{s},\widehat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\end{array}
\]
The first line of $ P(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
values of $h$ and $k$.
The second line accounts for the analogous case where $s$ is not provided by the context ($\link{\underline{s}_i}{\underline{s}_o}$). 
The condition $k\geq 1$ guarantees that $s$ will remain present (the continuation is $P_s$).
%Here taking
%$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
The third line accounts for the case where $s$ is tested for presence, but it is neither produced nor provided by  the context. Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence.
%Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
% permutations of the solid links; nevertheless if the reactions, and their enumeration, is known
% the code of $P_s$ can be drastically simplified.

\paragraph{Processes for reactions.}
We assume that each reaction $a$ is assigned a progressive number $j$.
The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\widetilde{c}_i}{\widetilde{c}_o}$ is requested for any $c\in P_j$).
The next case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
The process is recursive so that reactions can be applied at any step.
\[
\begin{array}{lcl@{\quad}l}
P_{aj}& \defeq & 
 ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\widetilde{c}_i}{\widetilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
&& + \ \\
&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
&&+\\
&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
\end{array}
\]

\noindent
We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $aj$.
Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using the channel $\mathit{cxt}$, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels). 
In the following there is an example explaining this schema.
%\[
%\begin{array}{lcl@{\qquad}l}
%P_{aj}& \defeq & 
% ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \link{}{}
% \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{r_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
%&& + \ \\
%&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainend{r_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
%&&+\\
%&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainend{r_{i+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
%\end{array}
%\]

\paragraph{Processes for contexts.}
The context can vary at each step $n$ and for each entity $s \in S$, the context must say if the entity is provided or not. Let us denote by $C_n$ the context at step $n$, i.e. the set of entities that are made available. Correspondingly, we introduce another process $\mathit{Cxt}^n$ defined as follows:
\[
\begin{array}{lcl}
\mathit{Cxt}^n &\defeq &
%\left \{
^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n} \,  \backslash_{p_1}. \mathit{Cxt}^{n+1}\\
%+ \\
% ^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in I2_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in Oj_n} \, \backslash_{p_1}
% \chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{p_1}. Cxt_{s}^{n+1} & \ \mbox{if the context provides $s$  at the $n$-th step}\\[5pt] 
% \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_s^{n+1} & \ \mbox{otherwise}
%\end{array}\right.\\
%Cxt_s & \defeq & Cxt_s^1\\
%\end{array}
%\right.\\
\end{array}
\]

%,  participating in each transition and saying if, the entity $s$ is provided by the context or not.
%As done for the reactions, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
%the chains formed by the application of all the contexts. 
%For each entity $s$ with number $j$, at step $n >0$  there are two possible behaviours:

@@

We only consider $\mathit{Cxt}^n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (if $s\in C_0$ the process $P_s$ will be present initially, otherwise $\overline{P_s}$ will be present).

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
$$
\llbracket{\cal A},\gamma \rrbracket 
\triangleq
 \restrict{\mathit{names}}(\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
$$
where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ctxs}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}$, with  $\mathit{reacts}$ be the set of reaction names $r_j $, $\mathit{cxts}$ the set of context names $\{\mathit{cxt}\}$,  $\mathit{ents}$ the set of decorated entity names  $\{s_i,s_o,\widehat{s}_i,\widehat{s}_o,\widetilde{s}_i,\widetilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $\mathit{prods}$ be the set of names for production $p_j$ associated to each reaction. 
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$  for $u$ the number of reacts.
\end{definition}

@@

It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes 
%$P_a$, with $a \in A$, and  $\mathit{Cxt}^n$ and  $P_s$, with $s \in S$,  
running in parallel to  interact in that transition.
This is due to the fact that  all the channels $r_j$, $p_j$, $\mathit{cxt}$, including those for decorated names $s_{i},s_{o},\overline{s}_i,\overline{s}_o$ are  restricted.
Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation. If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{\mathit{names}}\upsilon} P')$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$; 
  the channel $\mathit{cxt}$ appears in $\upsilon$;
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^n$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.

@@

With the next proposition, we analyse the structure of a c\CNA \ 
process encoding of  a reactive process after one step transition.
In the following four statements, for brevity, we let ${\cal A} = (S, A)$ be a RS, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with $\gamma=\{C_i\}_{i\in\mathbb{N}}$ and $\delta=\{D_i\}_{i\in\mathbb{N}}$. Moreover, we denote by $\pi^j$ the shift of $\pi$ starting at the $j$-th state sequence; formally we let $\pi^j=(\gamma^j,\delta^j)$ with
$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ with 
  $C_0'=C_j \cup D_j$, $D'_0 = \emptyset$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.

 \begin{restatable}[Correctness 1]{proposition}{propcorruno}
% \begin{proposition}(correctness 1)\\
 \label{prop:corr1}
% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
 Let 
 $P=\llbracket{\cal A},\gamma\rrbracket$ with
$$P  = \restrict{\mathit{names}} (\Pi_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \Pi_{s \in C_0}P_s ~|~ \Pi_{s \notin C_0}\overline{P}_s).
$$
 If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:
 \begin{enumerate}
 \item
  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$, and
  \item 
 $P'= \restrict{\mathit{names}} (\Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
 \end{enumerate}
Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
 %\end{proposition}
\end{restatable}

@@

With the following propositions, we prove that, given a RS ${\cal A} = (S, A)$ and an extended  interactive process $\pi=(\gamma,\delta)$, then the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of $\pi$.
